---
title: "Racial Biases in Police Stops: Analysis of the Stanford Open Policing Project"
author: "Stanley Lo"
subtitle: "May 2025"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

### Overview

- Analyze differences in arrest and search rates by race during police stops
- Focus on three Bay Area California cities:
  - San Francisco 
  - Oakland 
  - San Jose 
- Time Frame: 2014 to 2017.
- Data source: Stanford Open Policing Project.

---

### Arrest Rates By Race

SELECT 
    city,
    subject_race,
    COUNT(*) AS total_stops,
    AVG(CASE WHEN arrest_made THEN 1 ELSE 0 END) AS arrest_rate
FROM (
    SELECT 
        'San Francisco' AS city,
        YEAR(date) AS stop_year,
        subject_race,
        arrest_made
    FROM ca_san_francisco_2020_04_01
    WHERE YEAR(date) BETWEEN 2014 AND 2017
    UNION ALL
---

## Arrest Rates by Race (cont.)
```{r}
library(ggplot2)
  ggplot(arrest_data, aes(x = subject_race, y = arrest_rate, fill = city)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.6) +
  labs(title = "Arrest Rates by Race (2014-2017)", x = "Subject Race", y = "Arrest Rate", fill = "City")
```

---

## Arrest Rates by Race - Analysis

- Black individuals have the highest number of arrests out of all arrests in **every city**.
- Oakland shows the **largest disparity**:
  - Arrest rate for Black individuals nearly **twice** that of the next highest race.
- Hispanic individuals consistently have the **second highest** arrest rates.

---
### Search Rates By Race

SELECT 
    city,
    subject_race,
    COUNT(*) AS total_stops,
    ROUND(100.0 * AVG(CASE WHEN search_conducted THEN 1 ELSE 0 END), 2) AS search_rate
FROM (
    SELECT 
        'San Francisco' AS city,
        subject_race,
        search_conducted
    FROM ca_san_francisco_2020_04_01
    WHERE YEAR(date) BETWEEN 2014 AND 2017

    UNION ALL

---

## Search Rates by Race (cont.)
```{r}
library(ggplot2)
ggplot(search_data, aes(x = subject_race, y = search_rate, fill = city)) + geom_col(position = position_dodge(width = 0.9), width = 0.6) + labs(title = "Search Rates by Race (2014-2017)", x = "Subject Race", y = "Search Rate (%)", fill = "City")
```
---

## Search Rates by Race - Analysis

- Search rates are higher for **Black and Hispanic** individuals across all cities.
- In San Jose:
  - Hispanic search rates sometimes **exceed** those of Black individuals.
  - The gap between Black and Hispanic search rates **narrows**.
- In Oakland and San Francisco:
  - The disparity between Black and Hispanic search rates remains **significant**.

---

## Interpretation of Arrest and Search Rates

### What the Data Suggests

- Out of all arrests, Black and Hispanic people experience high rates of:
  - Arrests
  - Searches
- This pattern is consistent across major Bay Area and Southern California cities.
- Indicates **potential racial disparities** in policing practices if taking a look at proportion of races in each city.

---

### Insights

  - High arrest rates for Black and Hispanic individuals out of all arrests made.
  - High search rates for Black and Hispanic individuals out of all searches made.

---

## Conclusion

### Summary Points

- Potential racial disparities exist in arrest and search rates in San Francisco, Oakland, and San Jose.
- Black and Hispanic people make up most of the arrest and search rates.
- These disparities highlight potential challenges in achieving fair policing practices.

---

## Limitations

- It would've been helpful to collect data on the demographics of the populations in each city.

- In addition, looking at cities in Southern California or a larger city like New York City would have added to the argument for reforms if the search and arrest rates were similar or gone against the argument for reforms.
---

## Thank You for listening to me!

**Data Source:**  
Stanford Open Policing Project  
https://openpolicing.stanford.edu/data/

**Citation:**  
Pierson, Emma, Camelia Simoiu, Jan Overgoor, Sam Corbett-Davies, Daniel Jenson, Amy Shoemaker, Vignesh Ramachandran, et al. 2020.  
“A Large-Scale Analysis of Racial Disparities in Police Stops Across the United States.” *Nature Human Behaviour*, 1–10.
