---
title: "Who Asks the Most Questions in Friends?"
format: html
---

Questions of Interest: "How do questions change throughout the seasons of Friends, and which characters ask the most questions?"


```{r}
library(friends)
suppressPackageStartupMessages(library(tidyverse))
library(stringr)
library(ggplot2)
data(friends)

friends <- friends |>
mutate(word_count = str_count(text, "\\w+"), 
first_word = str_extract(text, "^\\w+"),
ends_with_question_mark = str_detect(text, "\\?$"))


question_words <- "Who|What|When|Where|Why|How"
friends <- friends |>
mutate(starts_with_question_word = str_detect(text,  str_c("^(", question_words, ")\\b")),
contains_question_word = str_detect(text, str_c("\\b(", question_words, ")\\b")),
question_word_count = str_count(text, str_c("\\b(", question_words, ")\\b")))

friends <- friends |>
mutate(complex_question = str_detect(text, str_c("^(?!(", question_words, ")\\b).*\\?$")))


season_questions <- friends |>
group_by(season) |>
summarise(question_ratio = mean(ends_with_question_mark),
complex_question_ratio = mean(complex_question))

```

```{r}

ggplot(season_questions, aes(x = season)) +
geom_line(aes(y = question_ratio, color = "All Questions")) +
geom_line(aes(y = complex_question_ratio, color = "Complex Questions")) + labs(title = "Question Usage Throughout Friends Seasons", x = "Season", y = "Proportion of Lines", color = "Question Type")

```

This graph shows how question usage is changing throughout the
multiple seasons of the show Friends. We see that all questions asked are stable but a slight dip in the proportion of lines dedicated to questions in the middle of the graph. At the same time, complex questions took a hit as well but since the proportional difference is only around 0.04, it is not a significant drop in questions being asked in the show Friends.

```{r}

main_characters <- c("Monica Geller", "Rachel Green", "Phoebe Buffay", "Chandler Bing", "Joey Tribbiani", "Ross Geller")

character_questions <- friends |>
filter(speaker %in% main_characters) |>
group_by(speaker) |>
summarise(question_ratio = mean(ends_with_question_mark),
question_word_ratio = mean(contains_question_word),
total_lines = n()) |>
arrange(desc(question_ratio))

```

```{r}
ggplot(character_questions, aes(x = speaker, y = question_ratio)) + geom_bar(stat = "identity", fill = "blue") + labs(title = "Proportion of Questions Asked by Main Characters", x = "Character", y = "Proportion of Lines Ending with Question Mark")
```

From our analysis of characters and the questions they ask.
We see that Monica seems to be the character that asks the most questions with around 24% of her lines being QUESTIONS 
which is insane. Besides that, the rest of the characters
also have similar rates of asking questions with everyone
except Phoebe having a proportion of lines ending with
question marks above 20%. Because Monica is asking the most
questions, we see that Monica is possibly also questioning
the decisions of her Friends much more than the others. This
could show the differentiated strategy in which the directors
are trying to develop the personalities of each character.

CITATIONS:

https://github.com/EmilHvitfeldt/friends

Character Mining Project: https://www.emorynlp.org/#